---
title: "ANCOM-BC2"
output: html_document
date: "2025-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
##DIFFERENTIAL ABUNDANCE ANALYSIS OF MICROBIAL FUNCTIONAL GENES USING ANCOM-BC2

```{r load libraries}
#BiocManager::install("mia")
#install.packages("openxlsx")
library(mia)
library(dplyr)
library(microbiome)
library(ggplot2)
library(openxlsx)
library(metagMisc)
library(ANCOMBC)
```
```{r load phyloseq objects}
#METABOLIC genes
Functional_genes_GLOM <- readRDS("~/Metagenomics/phyloseq/Functional_genes_physeq.rds")
Functional_genes_CAT <- tax_glom(Functional_genes_GLOM, taxrank = "Category") #merge at category level 
taxa_names(Functional_genes_CAT) <- tax_table(Functional_genes_CAT)[,1]
Functional_genes_FUNCTION <- tax_glom(Functional_genes_GLOM, taxrank = "Func_abb") #merge at function level
taxa_names(Functional_genes_FUNCTION) <- tax_table(Functional_genes_FUNCTION)[,2]

#CAZy genes
CAZy_ps <- readRDS("~/Metagenomics/phyloseq/CAZy_GLOM.rds")
CAZy_ps_EC <- tax_glom(CAZy_ps, taxrank = "Enzyme_Class")
taxa_names(CAZy_ps_EC) <- tax_table(CAZy_ps_EC)[,1]

#Bacterial GENERA
genus_ps <- readRDS("~/Metagenomics/Kraken/genus_ps.rds")
taxa_names(genus_ps) <- tax_table(genus_ps)[,1]
genus_ps_RA <- readRDS("~/Metagenomics/Kraken/genus_ps_RA.rds") #for filtering criteria
taxa_names(genus_ps_RA) <- tax_table(genus_ps_RA)[,1]
```
```{r}
#filter bacterial genera based on relative abundance
genus_ps_RA2 <- filter_taxa(genus_ps_RA, function(x) sum(x > 0.5) > (0.05*length(x)), TRUE) 
keep_taxa <- taxa_names(genus_ps_RA2)
genus_ps2 = subset_taxa(genus_ps, Phylum %in% keep_taxa)
```
```{r set Hirakimata forest as reference group}
sample_data(Functional_genes_CAT)$Forest <- factor(sample_data(Functional_genes_CAT)$Forest, levels = c("Hirakimata", "Puketi", "Glenfern", "Windy_Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara"))
levels(sample_data(Functional_genes_CAT)$Forest)

sample_data(Functional_genes_FUNCTION)$Forest <- factor(sample_data(Functional_genes_FUNCTION)$Forest, levels = c("Hirakimata", "Puketi", "Glenfern", "Windy_Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara"))
levels(sample_data(Functional_genes_FUNCTION)$Forest)

sample_data(Functional_genes_GLOM2)$Forest <- factor(sample_data(Functional_genes_GLOM2)$Forest, levels = c("Hirakimata", "Puketi", "Glenfern", "Windy_Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara"))
levels(sample_data(Functional_genes_GLOM2)$Forest)

sample_data(CAZy_ps_EC)$Forest <- factor(sample_data(CAZy_ps_EC)$Forest, levels = c("Hirakimata", "Puketi", "Glenfern", "Windy_Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara"))
levels(sample_data(CAZy_ps_EC)$Forest)

sample_data(CAZy_ps2)$Forest <- factor(sample_data(CAZy_ps2)$Forest, levels = c("Hirakimata", "Puketi", "Glenfern", "Windy_Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara"))
levels(sample_data(CAZy_ps2)$Forest)

sample_data(genus_ps2)$Forest <- factor(sample_data(genus_ps2)$Forest, levels = c("Hirakimata", "Puketi", "Glenfern", "Windy_Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara"))
levels(sample_data(genus_ps2)$Forest)
```
#reformat 'taxa' names to be compatible with ANCOM-BC analysis
```{r}
colnames(tax_table(Functional_genes_CAT)) <- c("Phylum", "Class", "Order", "Family", "Genus")
colnames(tax_table(Functional_genes_FUNCTION)) <- c("Phylum", "Class", "Order", "Family", "Genus")
colnames(tax_table(Functional_genes_GLOM2)) <- c("Phylum", "Class", "Order", "Family", "Genus")
colnames(tax_table(CAZy_ps_EC)) <- c("Phylum", "Class", "Order", "Family")
colnames(tax_table(CAZy_ps2)) <- c("Phylum", "Class", "Order", "Family")
colnames(tax_table(genus_ps2)) <- c("Phylum", "Class", "Order", "Family", "Genus", "Species","Genus.1")
```

##RUN ANCOMBC-BC2 ANALYSIS

##CATEGORY LEVEL - merge with cazy gene categories for ANCOM-BC analysis
```{r}
##check levels match
levels(sample_data(Functional_genes_GLOM2)$Forest)
levels(sample_data(CAZy_ps_EC)$Forest)
Functional_genes_GLOM.merged <- merge_phyloseq(Functional_genes_GLOM2, CAZy_ps_EC)
```
```{r}
set.seed(71)
ancombc_category = ancombc2(data = Functional_genes_GLOM.merged,
                  tax_level = "Phylum",
                  fix_formula = "Forest", 
                  p_adj_method = "holm", 
                  prv_cut = 0, lib_cut = 0, s0_perc = 0.05,
                  pseudo = 0.1, pseudo_sens = FALSE,
                  group = "Forest", struc_zero = FALSE, neg_lb = FALSE,
                  alpha = 0.05, 
                  global = TRUE, pairwise = TRUE,
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100))
```
##FUNCTIONAL LEVEL
```{r}
set.seed(71)
ancombc_function = ancombc2(data = Functional_genes_FUNCTION,
                  tax_level = "Class",
                  fix_formula = "Forest", 
                  p_adj_method = "holm", 
                  prv_cut = 0, lib_cut = 0, s0_perc = 0.05,
                  pseudo = 0.1, pseudo_sens = FALSE,
                  group = "Forest", struc_zero = FALSE, neg_lb = FALSE,
                  alpha = 0.05, 
                  global = TRUE, pairwise = TRUE,
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100))
```
##GENE LEVEL
```{r}
set.seed(71)
ancombc_genes = ancombc2(data = Functional_genes_GLOM2,
                  tax_level = "Genus",
                  fix_formula = "Forest", 
                  p_adj_method = "holm", 
                  prv_cut = 0, lib_cut = 0, s0_perc = 0.05,
                  pseudo = 0.1, pseudo_sens = FALSE,
                  group = "Forest", struc_zero = FALSE, neg_lb = FALSE,
                  alpha = 0.05, 
                  global = TRUE, pairwise = TRUE,
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100))
```
##CAZY GENES
```{r}
set.seed(71)
ancombc_CAZy = ancombc2(data = CAZy_ps2,
                  tax_level = "Class",
                  fix_formula = "Forest", 
                  p_adj_method = "holm", 
                  prv_cut = 0, lib_cut = 0, s0_perc = 0.05,
                  pseudo = 0.1, pseudo_sens = FALSE,
                  group = "Forest", struc_zero = FALSE, neg_lb = FALSE,
                  alpha = 0.05, 
                  global = TRUE, pairwise = TRUE,
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100))
```
##BACTERIAL GENERA
```{r}
set.seed(71)
ancombc_genus = ancombc2(data = genus_ps2,
                  tax_level = "Phylum",
                  fix_formula = "Forest", 
                  p_adj_method = "holm", 
                  prv_cut = 0, lib_cut = 0, s0_perc = 0.05,
                  pseudo = 0.1, pseudo_sens = FALSE,
                  group = "Forest", struc_zero = FALSE, neg_lb = FALSE,
                  alpha = 0.05, 
                  global = TRUE, pairwise = TRUE,
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100))
```
##Export result datasets
```{r}
ANCOM_CATEGORIES <- cbind(ancombc_category$res_global, ancombc_category$res_pair[,182:217])
ANCOM_GENES <- cbind(ancombc_genes$res_global, ancombc_genes$res_pair[,182:217])
ANCOM_FUNCTIONS <- cbind(ancombc_function$res_global, ancombc_function$res_pair[,182:217])
ANCOM_CAZY <- cbind(ancombc_CAZy$res_global, ancombc_CAZy$res_pair[,182:217])
ANCOM_BACTERIA <- cbind(ancombc_genus$res_global, ancombc_genus$res_pair[,182:217])
```
```{r identify variables with no significant pairwise differences}
No_PW_CAT <- ANCOM_CATEGORIES[apply(ANCOM_CATEGORIES[,6:41], 1, function(x) all(x==FALSE)),]
No_PW_GENES <- ANCOM_GENES[apply(ANCOM_GENES[,6:41], 1, function(x) all(x==FALSE)),]
No_PW_FUNCTIONS <- ANCOM_FUNCTIONS[apply(ANCOM_FUNCTIONS[,6:41], 1, function(x) all(x==FALSE)),]
No_PW_CAZY <- ANCOM_CAZY[apply(ANCOM_CAZY[,6:41], 1, function(x) all(x==FALSE)),]
No_PW_BACTERIA <- ANCOM_BACTERIA[apply(ANCOM_BACTERIA[,6:41], 1, function(x) all(x==FALSE)),]
```
```{r}
list_of_datasets <- list("Gene"= ANCOM_GENES, "Function"= ANCOM_FUNCTIONS, "Category"= ANCOM_CATEGORIES, "CAZy"= ANCOM_CAZY, "Bacteria"= ANCOM_BACTERIA)
write.xlsx(list_of_datasets, file = "~/Metagenomics/phyloseq/Results/ANCOMBC.xlsx")
```
##PLOTTING RESULTS
##CATEGORY
```{r}
category_abundances <- as.data.frame(ancombc_category$bias_correct_log_table)
Forests <- as.data.frame(sample_data(Functional_genes_GLOM.merged)$Forest)
colnames(Forests) <- "Forest"
category_abundances_DF <- cbind(Forests, t(category_abundances))

##extract W-value
category_W_value <- ANCOM_CATEGORIES[,1:2]
category_W_value2 <- cbind(rep(category_W_value[,1], each=9), rep(category_W_value[,2], each=9))

##melt dataframe
category_abundances_m <- melt(category_abundances_DF)
category_abundances_sum <- category_abundances_m %>%
  dplyr::group_by(Forest, variable) %>%
  dplyr::summarize(Abundance = mean(value), SD = sd(value), n = n(), SE = SD/sqrt(n))

#add gene metadata
category_abundances_sum <- category_abundances_sum[order(category_abundances_sum$variable),]
category_abundances_sum <- cbind(category_W_value2, category_abundances_sum)
colnames(category_abundances_sum) <- c("Category", "W_value","Forest", "variable",  "Abundance", "SD", "n", "SE")
```
```{r remove categories that had no significant pairwise differences}
category_abundances_sum2 <- category_abundances_sum[! category_abundances_sum$Category %in% No_PW_CAT$taxon , ]
```
```{r EDIT VARIABLE NAMES}
levels(category_abundances_sum2$Forest) <- c("Hirakimata", "Puketi", "Glenfern", "Windy Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara")
category_abundances_sum2$W_value <- as.numeric(category_abundances_sum2$W_value)
levels(category_abundances_sum2$variable) <- c("Amino acid metabolism", "Aromatics degradation", "Arsenate cycling", "Auxiliary Activities [CAZy]",  "Carbohydrate Binding Module [CAZy]", "Carbohydrate Esterase [CAZy]",  "Carbon degradation",  "Carbon metabolism", "Chlorite reduction", "Fatty acid degradation", "Fermentation", "Glycoside Hydrolase [CAZy]", "GlycosylTransferase [CAZy]", "AOX degradation", "Hydrogen metabolism",  "Iron cycling", "Manganese cycling", "C1 metabolism", "Nitrogen cycling", "Polysaccharide Lyase [CAZy]", "Sulfur metabolism")
```
##PLOT
```{r plot category abundances, warning=FALSE}
Category_plot <- ggplot(category_abundances_sum2, aes(x = Abundance, y = reorder(variable, +W_value), fill = Forest)) +
  geom_point(aes(size = W_value), color = "black", alpha = 1, shape = 21) +  # Black outline around points
  geom_errorbar(aes(xmin = Abundance - SE, xmax = Abundance + SE, color = Forest), size = 1, width = 0.3) + #Colored by Forest
  scale_fill_brewer(palette = "YlOrRd") +  # Fill color for points
  scale_color_brewer(palette = "YlOrRd") +  # Color scale for error bars
  scale_size_continuous(range = c(2, 10), guide = "none") + 
  theme_bw(base_size = 20) +
  xlab("Bias-corrected log abundance") +
  theme(axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black", size = 15),
    axis.title.x = element_text(colour = "black", size = 15),
    axis.title.y = element_text(colour = "black", size = 15),
    legend.position = "right") +
  ylab("") +
  labs(fill = "Forest", color = "Forest") 
Category_plot <- Category_plot + guides(fill = guide_legend(override.aes = list(size=6)))
```
```{r export plot}
png('~/Metagenomics/Manuscript/Figures/METABOLIC_categories.png', width = 6000, height = 5000, res = 500)
Category_plot
```
##FUNCTION
```{r}
function_abundances <- as.data.frame(ancombc_function$bias_correct_log_table)
function_abundances_DF <- cbind(Forests, t(function_abundances))

##extract W-value
function_category <- as.data.frame(tax_table(Functional_genes_FUNCTION)[,1])
function_category <- as.data.frame(function_category[order(function_category$Phylum),])

function_W_value <- ANCOM_FUNCTIONS[,1:2]
function_W_value2 <- cbind(rep(function_category[,1], each=9),rep(function_W_value[,1], each=9), rep(function_W_value[,2], each=9))

##melt dataframe
function_abundances_m <- melt(function_abundances_DF)
function_abundances_sum <- function_abundances_m %>%
  dplyr::group_by(Forest, variable) %>%
  dplyr::summarize(Abundance = mean(value), SD = sd(value), n = n(), SE = SD/sqrt(n))

#add gene metadata
function_abundances_sum <- function_abundances_sum[order(function_abundances_sum$variable),]
function_abundances_sum <- cbind(function_W_value2, function_abundances_sum)
colnames(function_abundances_sum) <- c("Category", "Function", "W_value","Forest", "variable",  "Abundance", "SD", "n", "SE")
```
```{r}
levels(function_abundances_sum$Forest) <- c("Hirakimata", "Puketi", "Glenfern", "Windy Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara")
function_abundances_sum$W_value <- as.numeric(function_abundances_sum$W_value)
```
```{r remove categories with only one level i.e. already been plotted}
function_abundances_sum$Category <- as.factor(function_abundances_sum$Category)
remove <- c("Halogenated compound degradation", "Aromatics degradation", "Arsenate cycling ", "Chlorite reduction", "Manganese oxidation", "Selenate reduction", "Fatty acid degradation", "Manganese cycling")
function_abundances_sum <- function_abundances_sum[!function_abundances_sum$Category %in% remove , ]
```
```{r plot function abundances, warning=FALSE}
function_plot <- ggplot(function_abundances_sum, aes(x = Abundance, y = reorder(variable, +W_value), fill = Forest)) +
  geom_point(aes(size = W_value), color = "black", alpha = 1, shape = 21) +  # Black outline around points
  geom_errorbar(aes(xmin = Abundance - SE, xmax = Abundance + SE, color = Forest), size = 1, width = 0.3) + #Colored by Forest
  scale_fill_brewer(palette = "YlOrRd") +  # Fill color for points
  scale_color_brewer(palette = "YlOrRd") +  # Color scale for error bars
  scale_size_continuous(range = c(2, 10), guide = "none") + 
  theme_bw(base_size = 20) +
  xlab("Bias-corrected log abundance") +
  theme(axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black", size = 15),
    axis.title.x = element_text(colour = "black", size = 15),
    axis.title.y = element_text(colour = "black", size = 15),
    legend.position = "right") +
  ylab("") +
  labs(fill = "Forest", color = "Forest") 
function_plot <- function_plot + guides(fill = guide_legend(override.aes = list(size=6)))
```
```{r}
function_plot2 <- function_plot + facet_grid(rows = vars(Category), scales = "free", space = "free")+
  theme(strip.background = element_blank(), strip.text = element_blank())
```
```{r}
png('~/Metagenomics/Manuscript/Figures/METABOLIC_functions.png', width = 7500, height = 8000, res = 500)
function_plot2
```
##GENES
```{r}
gene_abundances <- as.data.frame(ancombc_genes$bias_correct_log_table)
gene_abundances_DF <- cbind(Forests, t(gene_abundances))

##extract W-value
gene_metadata <- as.data.frame(tax_table(Functional_genes_GLOM2)[,1])
gene_metadata2 <- as.data.frame(taxa_names(Functional_genes_GLOM2))
gene_metadata3 <- as.data.frame(tax_table(Functional_genes_GLOM2)[,2])
gene_W_value <- ANCOM_GENES[,1:2]
gene_W_value2 <- cbind(rep(gene_metadata[,1], each=9), rep(gene_metadata2[,1], each=9), rep(gene_metadata3[,1], each=9), rep(gene_W_value[,2], each=9))

##melt dataframe
gene_abundances_m <- melt(gene_abundances_DF)
gene_abundances_sum <- gene_abundances_m %>%
  dplyr::group_by(Forest, variable) %>%
  dplyr::summarize(Abundance = mean(value, na.rm=FALSE), SD = sd(value), n = n(), SE = SD/sqrt(n))

#add gene metadata
gene_abundances_sum <- gene_abundances_sum[order(gene_abundances_sum$variable),]
gene_abundances_sum <- cbind(gene_W_value2, gene_abundances_sum)
colnames(gene_abundances_sum) <- c("Category","gene", "Function","W_value","Forest", "variable",  "Abundance", "SD", "n", "SE")
```
```{r}
levels(gene_abundances_sum$Forest) <- c("Hirakimata", "Puketi", "Glenfern", "Windy Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara")
gene_abundances_sum$W_value <- as.numeric(gene_abundances_sum$W_value)
```

##PLOT GENES AND SPLIT DATAFRAME ACCORDING TO GENE CATEGORY
```{r}
#split dataframe
abundance_genes.split <- split(gene_abundances_sum, ~ Category)

#remove variables not of interest
keep_variables <- c("Carbon degradation", "Carbon metabolism", "Methane metabolism", "Nitrogen cycling", "Sulfur metabolism")
abundance_genes.split2 <- abundance_genes.split[names(abundance_genes.split) %in% keep_variables == TRUE]
```
```{r plot genes, warning=FALSE}
genes_plots <- lapply(names(abundance_genes.split2), function(name) {
  x <- abundance_genes.split2[[name]]
  ggplot(x, aes(x = Abundance, y = reorder(variable, +W_value), fill = Forest)) +
    geom_point(aes(size = W_value), color = "black", alpha = 1, shape = 21) +
    geom_errorbar(aes(xmin = Abundance - SE, xmax = Abundance + SE, color = Forest), size = 1, width = 0.3) +
    scale_fill_brewer(palette = "YlOrRd") +
    scale_color_brewer(palette = "YlOrRd") +
    scale_size_continuous(range = c(2, 10), guide = "none") + 
    theme_bw(base_size = 20) +
    xlab("Bias-adjusted log abundance") +
    theme(axis.text.x = element_text(colour = "black", size = 15),
          axis.text.y = element_text(colour = "black"),
          axis.title.x = element_text(colour = "black"),
          axis.title.y = element_text(colour = "black"),
          legend.position = "right",
          plot.title = element_text(size = 20)) +
    ylab("") +
    labs(fill = "Forest", color = "Forest") +
    #scale_y_discrete(limits = rev) +
    ggtitle(name) +  # Use the list name as the title
    guides(fill = guide_legend(override.aes = list(size = 6)))
})
```
```{r combine C and N genes and export}
C_plot <- genes_plots[[1]] + labs(title = "A) C degradation genes") + theme(axis.title.x = element_blank())
N_plot <- genes_plots[[4]] + labs(title = "B) N cycling genes")
CN_plot <- ggarrange(C_plot, N_plot, common.legend = TRUE, legend = "right", align = "v", ncol = 1, heights = c(2,1.2))
```
```{r export C and N genes}
png('~/Metagenomics/Manuscript/Figures/CN_genes.png', width = 6500, height = 8000, res = 500)
CN_plot
```
##export C metabolism, S metabolism, and methane metabolism plots for supplementary materials
```{r}
S_plot <- genes_plots[[5]] + labs(title = "Sulfur metabolism genes")
png('~/Metagenomics/Manuscript/Figures/S_plot.png', width = 5000, height = 6000, res = 500)
S_plot
```
```{r}
Methane_plot <- genes_plots[[3]] + labs(title = "Methane metabolism genes")
png('~/Metagenomics/Manuscript/Figures/Methane_plot.png', width = 4500, height = 4000, res = 500)
Methane_plot
```
```{r redo C met plot to filter genes with low W values}
C_met_plot <- ggplot(subset(abundance_genes.split2$`Carbon metabolism`, W_value > 70), aes(x = Abundance, y = reorder(variable, +W_value), fill = Forest)) +
  geom_point(aes(size = W_value), color = "black", alpha = 1, shape = 21) +  # Black outline around points
  geom_errorbar(aes(xmin = Abundance - SE, xmax = Abundance + SE, color = Forest), size = 1, width = 0.3) + #Colored by Forest
  scale_fill_brewer(palette = "YlOrRd") +  # Fill color for points
  scale_color_brewer(palette = "YlOrRd") +  # Color scale for error bars
  scale_size_continuous(range = c(2, 10), guide = "none") + 
  theme_bw(base_size = 20) +
  xlab("Bias-corrected log abundance") +
  theme(axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black", size = 15),
    axis.title.x = element_text(colour = "black", size = 15),
    axis.title.y = element_text(colour = "black", size = 15),
    legend.position = "right") +
  ylab("") +
  labs(fill = "Forest", color = "Forest") 
C_met_plot <- C_met_plot + guides(fill = guide_legend(override.aes = list(size=6)))+ labs(title = "Carbon metabolism genes")
png('~/Metagenomics/Manuscript/Figures/C_met_plot.png', width = 7500, height = 8500, res = 500)
C_met_plot
```
##CAZY GENE FAMILIES
```{r}
cazy_abundances <- as.data.frame(ancombc_CAZy$bias_correct_log_table)
cazy_abundances_DF <- cbind(Forests, t(cazy_abundances))

##extract W-value
cazy_EC.df <- as.data.frame(tax_table(CAZy_ps2)[,1])
cazy_EC.df <- as.data.frame(cazy_EC.df[order(cazy_EC.df$Phylum),])

cazy_W_value <- ANCOM_CAZY[,1:2]
cazy_W_value2 <- cbind(rep(cazy_EC.df[,1], each=9),rep(cazy_W_value[,1], each=9), rep(cazy_W_value[,2], each=9))

##melt dataframe
cazy_abundances_m <- melt(cazy_abundances_DF)
cazy_abundances_sum <- cazy_abundances_m %>%
  dplyr::group_by(Forest, variable) %>%
  dplyr::summarize(Abundance = mean(value), SD = sd(value), n = n(), SE = SD/sqrt(n))

#add gene metadata
cazy_abundances_sum <- cazy_abundances_sum[order(cazy_abundances_sum$variable),]
cazy_abundances_sum <- cbind(cazy_W_value2, cazy_abundances_sum)
colnames(cazy_abundances_sum) <- c("Enzyme Class", "Gene Family", "W_value","Forest", "variable",  "Abundance", "SD", "n", "SE")
```
```{r}
levels(cazy_abundances_sum$Forest) <- c("Hirakimata", "Puketi", "Glenfern", "Windy Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara")
cazy_abundances_sum$W_value <- as.numeric(cazy_abundances_sum$W_value)
```
```{r plot cazy abundances, warning=FALSE}
cazy_plot <- ggplot(subset(cazy_abundances_sum, W_value  > 45), aes(x = Abundance, y = reorder(variable, +W_value), fill = Forest)) +
  geom_point(aes(size = W_value), color = "black", alpha = 1, shape = 21) +  # Black outline around points
  geom_errorbar(aes(xmin = Abundance - SE, xmax = Abundance + SE, color = Forest), size = 1, width = 0.3) + #Colored by Forest
  scale_fill_brewer(palette = "YlOrRd") +  # Fill color for points
  scale_color_brewer(palette = "YlOrRd") +  # Color scale for error bars
  scale_size_continuous(range = c(2, 10), guide = "none") + 
  theme_bw(base_size = 20) +
  xlab("Bias-corrected log abundance") +
  theme(axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black", size = 15),
    axis.title.x = element_text(colour = "black", size = 15),
    axis.title.y = element_text(colour = "black", size = 15),
    legend.position = "right") +
  ylab("") +
  labs(fill = "Forest", color = "Forest") 
cazy_plot <- cazy_plot + guides(fill = guide_legend(override.aes = list(size=6)))
```
```{r}
cazy_plot2 <- cazy_plot + facet_grid(rows = vars(`Enzyme Class`), scales = "free", space = "free")+
  theme(strip.background = element_blank(), strip.text = element_blank())
```
```{r}
png('~/Metagenomics/Manuscript/Figures/METABOLIC_cazys.png', width = 7500, height = 8500, res = 500)
cazy_plot2
```
##BACTERIAL GENERA
```{r import dataset of bacterial orders, classes and families}
bacterial_taxonomy <- read.csv("~/Metagenomics/Kraken/bacterial_genera_taxonomy.csv")
bacterial_taxonomy_DF <- cbind(rep(bacterial_taxonomy[,1], each=9), rep(bacterial_taxonomy[,2], each=9), rep(bacterial_taxonomy[,3], each=9))
```
```{r}
genus_abundances <- as.data.frame(ancombc_genus$bias_correct_log_table)
genus_abundances_DF <- cbind(Forests, t(genus_abundances))

##extract W-value
genus_W_value <- ANCOM_BACTERIA[,1:2]
genus_W_value2 <- cbind(rep(genus_W_value[,1], each=9), rep(genus_W_value[,2], each=9))
genus_W_value2 <- cbind(bacterial_taxonomy_DF, genus_W_value2)

##melt dataframe
genus_abundances_m <- melt(genus_abundances_DF)
genus_abundances_sum <- genus_abundances_m %>%
  dplyr::group_by(Forest, variable) %>%
  dplyr::summarize(Abundance = mean(value), SD = sd(value), n = n(), SE = SD/sqrt(n))

#add gene metadata
genus_abundances_sum <- genus_abundances_sum[order(genus_abundances_sum$variable),]
genus_abundances_sum <- cbind(genus_W_value2, genus_abundances_sum)
colnames(genus_abundances_sum) <- c("Class", "Order","Family", "genus", "W_value","Forest", "variable",  "Abundance", "SD", "n", "SE")
```
```{r}
levels(genus_abundances_sum$Forest) <- c("Hirakimata", "Puketi", "Glenfern", "Windy Hill", "Laingholm", "Silverdale", "Gadgil",  "Oratia", "Kaiaraara")
genus_abundances_sum$W_value <- as.numeric(genus_abundances_sum$W_value)
genus_abundances_sum$Class <- as.factor(genus_abundances_sum$Class)
```
```{r plot genus abundances, warning=FALSE}
genus_plot <- ggplot(genus_abundances_sum, aes(x = Abundance, y = reorder(variable, +W_value), fill = Forest)) +
  geom_point(aes(size = W_value), color = "black", alpha = 1, shape = 21) +  # Black outline around points
  geom_errorbar(aes(xmin = Abundance - SE, xmax = Abundance + SE, color = Forest), size = 1, width = 0.3) + #Colored by Forest
  scale_fill_brewer(palette = "YlOrRd") +  # Fill color for points
  scale_color_brewer(palette = "YlOrRd") +  # Color scale for error bars
  scale_size_continuous(range = c(2, 10), guide = "none") + 
  theme_bw(base_size = 20) +
  xlab("Bias-corrected log abundance") +
  theme(axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black", size = 12),
    axis.title.x = element_text(colour = "black", size = 15),
    axis.title.y = element_text(colour = "black", size = 15),
    legend.position = "right") +
  ylab("") +
  labs(fill = "Forest", color = "Forest") +
  ggforce::facet_col(facets = vars(Class), scales = "free_y", space = "free") +
  theme(strip.text.y = element_text(angle=0), strip.text = element_text(size = 12))
genus_plot <- genus_plot + guides(fill = guide_legend(override.aes = list(size=6)))
```
```{r}
png('~/Metagenomics/Manuscript/Figures/Bacterial_genera.png', width = 6500, height = 8000, res = 500)
genus_plot
```
